6.878 Final Report

### 2013-12-09 report/paper.tex
more stuff for report

\subsection{Personal Genomics}
With the advent of efficient sequencing, the possibility of personalized medicine based on an inidividual's specific genetic composition and molecular phenotype becomes

\subsection{Snyder Dataset}
Snyder

Chen, et al. examined how to generate an integrative personal omics profile to examine how biological components change during healthy and diseased states and how to use this information to estimate disease risk and learn about diseased states. While

\subsection{Other time series}

@@

@@

Additionally, we used \verb!DAVID! Bioinformatics Database web service to find functional groups within each cluster, as a means of describing each cluster, with $kappa=70$ and $linkage = 0.50$. We were able to generate functional groups for several clusters.

\subsection{Visualization}
In order to make expression trend analysis easier from a clinical standpoint, we created a browser-based visualization tool using \verb!d3! that provides normalized/denormalized expression trends by cluster, dendrograms for cluster annotation, and 
\section{Future Goals}

\section{Commetary on experience}
Iris: I learned a lot on how to pull data and to compare results with the ones in a paper. I also learned a lot about

@@

>>>>

\subsection{Personal Genomics}
With the advent of efficient sequencing techniques and improved computation resources, the possibility of personalized medicine based on an inidividual's specific genetic composition and molecular phenotype has increased dramatically over the past few years.

\subsection{Snyder Dataset}
The original dataset consisted of a genomic profile generated from blood samples taken from a 54-year-old male at 20 time points over the course of $401$ days. The supplemented the genomic data with blood glucose sampling, and increased the sampling rate during a human rhinovirus infection and a respiratory syncytial virus infection.

Chen, et al. examined how to generate an integrative personal omics profile to examine how biological components change during healthy and diseased states and how to use this information to estimate disease risk and learn about diseased states. Their analysis included whole genome sequencing, complete transcriptome analysis, proteomic and metabolomic analyses, and autoantibody profiles.
Their multi-dimensional view of the individual showed how traits associated with disease could be monitored to identify physiological states.
In particular, they processed the various omics sets using spectral analysis, clustering, and pathway enrichment analysis.

\subsection{Other time series}
In theory, our algorithms can work on any expression time series.

We also tested our process on other pre-processed time series for gene expression including \verb!GSE675!, for HCMV infection of foreskin fibroblasts.

@@

\begin{center}
\begin{tabular}{c | ccc}
  & R & D & S\\ \hline

  R &  3 & -1 & -3\\
  D & -1 & 3 & -3\\
  S & -3 & -3 & 1\\
\end{tabular}
\quad
\begin{tabular}{c | ccc}
  & R & D & S\\ \hline

  R &  3 & -3 & -1\\
  D & -3 & 3 & -1\\
  S & -1 & -1 & 3\\
\end{tabular}
\quad
\begin{tabular}{c | ccc}
  & R & D & S\\ \hline

  R &  3 & -3 & -3\\
  D & -3 & 3 & -3\\
  S & -3 & -3 & 3\\
\end{tabular}
\end{center}

@@

Additionally, we used \verb!DAVID! Bioinformatics Database web service to find functional groups within each cluster, as a means of describing each cluster, with $kappa=70$ and $linkage = 0.50$. We were able to generate functional groups for several clusters.
The \verb!DAVID! clustering provided functional groups within each cluster, whereas corrected Fischer's Exact test tries to summarize an entire cluster. Since we were able to find significant groups within each cluster, but not for an entire cluster, this may indicate that our clustering is too general or that the clusters are too large.

\begin{figure}[h]
\includegraphics[height=3in]{dendrogram.png}
\caption{Dendrogram describing a cluster using our visualization tool.}
\label{fig:dendrogram}
\end{figure}

\subsection{Visualization}
In order to make expression trend analysis easier from a clinical standpoint, we created a browser-based visualization tool using \verb!d3! that provides normalized/denormalized expression trends by cluster, dendrograms for cluster annotation, and word clouds for a text summary of \verb!DAVID! gene groups by cluster.
\section{Future Goals}

We could integrate this with the SMART Genomics API to make our tool usable for all time series of gene expression.
\section{Commetary on experience}
Iris: I learned a lot on how to pull data and to compare results with the ones in a paper. I also learned a lot about

@@

\bibitem{GSE675} Browne EP, Wing B, Coleman D, Shenk T. Altered cellular mRNA levels in human cytomegalovirus-infected fibroblasts: viral block to the accumulation of antiviral mRNAs. J Virol 2001 Dec;75(24):12319-30. PMID: 11711622
